---
title: "K Nearest Neighbors"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

The K Nearest Neighbors (KNN) algorithm predicts the labels of the test dataset by looking at the labels of its closest neighbors in the feature space of the training dataset. The following example will use it for a classification problem.

## Dataset

The dataset can be found [here](https://www.openintro.org/data/index.php?data=loans_full_schema). It consists of 10,000 loans, and we will find whether a loan will be paid back based on the customer’s data.

```{r}
suppressPackageStartupMessages(library(tidyverse))

df <- read_csv('data/loans.csv', show_col_types = FALSE)
df <- subset(df, select = -c(loan_purpose))
head(df)
```

We will remove columns where there are a lot of missing (NA) or null values

```{r}
df %>%
  summarize_all(function(x){ sum(is.na(x)) })

### Remove NA columns
df <- df %>%
  select(-c(emp_title, emp_length, annual_income_joint, debt_to_income, verification_income_joint, debt_to_income_joint, months_since_last_delinq, months_since_90d_late, months_since_last_credit_inquiry, num_accounts_120d_past_due))

### Create column to determine if fully paid or not
df$fully_paid <- ifelse(df$loan_status == 'Fully Paid', 1, 0)

### Remove non-numeric columns
df <- df %>% select_if(is.numeric)

### Change fully paid to a factor
df$fully_paid <- as.factor(df$fully_paid)
```

## Training and Test data

Now we split the data into test and training datasets

```{r}
suppressPackageStartupMessages(library(caret))
set.seed(1000)

trainIndex <- createDataPartition(df$fully_paid, 
                                  times=1, 
                                  p = .8, 
                                  list = FALSE)
train <- df[trainIndex, ]
test <- df[-trainIndex, ]
```

## Feature scaling

We will now scale both the training and testing sets

```{r}
preProcValues <- preProcess(train, method = c("center", "scale"))
trainTransformed <- predict(preProcValues, train)
testTransformed <- predict(preProcValues, test)
```

## Model

Now we can construct the model

```{r}
knnModel <- train(
  fully_paid ~ ., 
  data = trainTransformed, 
	method = "knn", 
	trControl = trainControl(method = "cv"), 
	tuneGrid = data.frame(k = c(3,5,7))
  )
```

And choose the best k value

```{r}
best_model<- knn3(fully_paid ~ .,
                  data = trainTransformed,
                  k = knnModel$bestTune$k)
```

## Model evaluation

Now we can evaluate the model using metrics given by the `caret` package

```{r}
predictions <- predict(best_model, testTransformed, type = "class")
# Calculate confusion matrix
cm <- confusionMatrix(predictions, testTransformed$fully_paid)
print(cm)
```

Our model is very good, with a high Balanced Accuracy value

Sensitivity measures the ratio of actual positive cases that are correctly identified and our model has very high sensitivity

Specificity represents the ability of the model to correctly identify negative cases. Our model is less good at correctly identifying negative cases